---
title: <span style="color:cadetblue">Data Clinic Day 1</span>
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
##<span style="color:cadetblue">What is R?</span>

R is a language and environment for statistical computing and graphics

##<span style="color:cadetblue">Why R?</span>
*  It's free!
*  Popularity -- we can show published papers with plots
*  Awesome power
     * can handle complex and large data
     * can easily program complex simulations
     * can be used on High Performance Computer Clusters
     * supports multicore task distribution
*  Flexibility
*  Great support
    * StackOverflow
    * R-Bloggers
    
##<span style="color:cadetblue">CRAN - The Comprehensive R Archive Network</span> 
**CRAN**^[https://cran.r-project.org] is a network of ftp and web servers that store identical, up-to-date, versions of code and documentation and packages for R.

##<span style="color:cadetblue">Installing packages</span>

```
install.packages("<the package's name>")

library("<the package's name>")
```

##<span style="color:cadetblue">R Studio</span>
RStudio^[https://www.rstudio.com/products/RStudio/#Desktop] is an integrated development environment (IDE) for R. It includes a console, syntax-highlighting editor that supports direct code execution, as well as tools for plotting, history, debugging and workspace management. 

##<span style="color:cadetblue">The console</span>

Type anything at the prompt press ```enter```  , and R will evaluate it and print the answer.
```{r echo=TRUE}
1+1
```


##<span style="color:cadetblue">The script</span>

The first time you open **RStudio**, you will see three windows. A forth window is hidden by default, but can be opened by clicking the **File** drop-down menu, then **New File**, and then **R Script**.

Type anything at the prompt press ```Run```  , and R will evaluate it and print the answer (the same as before).

You can save this script to use it later-- show how

###<span style="color:cadetblue">Commenting on script</span>

```{r echo=TRUE}
1+1 # Anything after "#" will not be evaluated by R
```


##<span style="color:cadetblue">Arithmetics</span>
```{r echo=TRUE}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2^5

# Modulo
26%%3

```

### <span style="color:orangered">Exercise 1</span>
```
Repeat all the operations with different values
```

##<span style="color:cadetblue">Variable Assignment</span>
```{r echo=TRUE}
# Assign the value 12 to x
x <- 12

# Print out the value of the variable x
x
```

### <span style="color:orangered">Exercise 2</span>
```
1. Choose 2 different variable names 
2. Assign them each a value
3. Add them together
4. Create a new variable and assign it the addition

```
##<span style="color:cadetblue">Relational operators</span>

Operator    | Description
----------- | -------------
>           | Checks if 1st element is greater than the 2nd element 
<           | Checks if 1st element is less than the 2nd element
==          | Checks if 1st element is equal to the 2nd element
!=          | Checks if 1st element is unequal to the 2nd element

### <span style="color:orangered">Exercise 3</span>
```
1. Choose 2 different variable names 
2. Assign them each a value
3. Use the operators
```



