---
title: ""
author: ""
date: "`r Sys.Date()`"
output: html_document
---
#<span style="color:cadetblue">Data Types</span>

***
This module should include 
logic statements
syntax
data structure
assigning objects
troubleshooting
reading in data

```{r settings}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```


##<span style="color:cadetblue"> `R` Objects </span>

Everything in `R` is an object. But what does this mean to someone who doesn't have a computer science background? 

`R` doesn't let users directly access the computer's memory, instead users access many different specialized data structures. You can think of this design as a filing cabinet. One drawer contains textbooks, another pictures, another data tables, and so on. In this analogy, each drawer is a different object class. 

If we look in the textbook drawer, each book or object has a similar structure (front cover, title page, table of contents, etc.). Just like each book has a unique title, `R` objects have user assigned names. 

Objects in `R` are often assigned a name by using an **arrow** made from a the **less than** sign and **dash** <-

```{r make an object}
myname <- "Odum" #this line assigns "Odum" to an object 'myname'
myname #print the content of 'myname' object by running it in the console
```

The information in the object is retrived by calling the object's name. 


## <span style="color:cadetblue">Data Types</span>

One of the first concepts we'll need to understand how `R` organizes information. This is **_very important_** because nearly every task requires manipulating data. `R` has 5 basic ('*atomic*') classes :

  * logical (`TRUE`, `FALSE`)
  * character ("blue", "purple")
  * integer (1,2,3) *each number must be followed by `L`
  * numeric (1,1.5,2)
  * complex (i)

 <span style="color:orangered">**_Concept Check_**:</span> What is the class of `myname`?

  Based on the decsription above, we can guess that is should be a character. We can use a few base `R` functions to get information about objects. Some helpful functions include:

  * typeof() 
  * length()
  * attributes()
  
 <span style="color:orangered">**Practice**:</span> Verify `myname` class type and determine the length. You may need to pull up the documentation page for the functions. 

```{r character class}
#What is the class
?typeof() #look up documentation for typeof()
typeof(myname) #return the object type of "myname"

#How long is the object
?length()
length(myname)
```

This object has a length of one. Why?

Let's make an numeric class object with a length greater than 1

```{r numeric class}
y <- c(1:10) #colon indicates '1 to 10' 
typeof(y)
length(y)
attributes(y)
```

 <span style="color:orangered">**_Concept Check_**:</span>  How could we make `myname` longer than 1?

## <span style="color:cadetblue">Data Structure</span>

These 5 classes or modes can then be combind to create many different types of *data structure*. Common data structures include:

  * vector
  * matrix 
  * list 
  * factors 
  * data frame
  
These data structures can consist of either a single class or a mixture of data types. Let's look at examples of each type. 

###<span style="color:cadetblue">Vectors</span>

```Vectors``` are the most common data structure. Vectors (the `R` data ) are vectors of elements can be character, logical, numeric or integer class.  
```{r Vectors}
#concatenating objects into a vector using "c"
character_vector <- c("a", "b", "c")
numeric_vector <- c(1, 10, 49)
boolean_vector <- c("TRUE", "FALSE", "TRUE")
```

You can creating ```Vectors``` using handy functions
```{r Vectors1}
vec1 <-  rep(1:4, 2)
vec2 <- rep(1:4, each = 2) 
vec3<-seq(1, 15, by = 2)
```

You can transform and use operators to create other ```vectors```. Check the different between these two vectors
```{r Vectors2}
vec4<-vec1+vec2+vec3

vec5<-sum(vec1,vec2,vec3)

```

Accessing specific specific ```elements``` in the ```vector```
```{r Vectors3}
vec4[1] # first elements

vec4[1:3] # first three elements

vec4[c(1,2,4)] # first second and fourth elements

vec6 <- c(1:20) #creating a vector 
x <- c(3, 5, 12) # creating another vector 

vec6[x]# the elements of vec6 that coicinde with vector x

vec6[-x] # What happened here?


vec7 <- c(1:5,2,2,2,2,2) #creating a vector 

vec7[vec7 > 4] # elements bigger than 4

vec7[vec7 < 4] # elements less than 4

which(vec7 < 4) # position

which(vec7 == 2) # position

```
###<span style="color:cadetblue">Matrices</span>

```Matrices``` are a special type of vector. The data is organized into columns and rows giving the matrix dimensions. There are many ways to make a matrix, but the default is to fill column-wise. 
```{r Matrices}
rnames <- c("R1", "R2")
cnames <- c("C1", "C2")
m<-matrix(data = NA, nrow = 2, ncol = 2, byrow=TRUE,
  dimnames=list(rnames, cnames))
```

<span style="color:orangered">**Practice**:</span> Add more rows and columns to the **matrix** and fill it with 1s and change the names of rows and columns
```{r Matrices1}
m<-matrix(data = , nrow = , ncol = , byrow = TRUE, dimnames = list(rnames, cnames))
```

Another way to build a matrix is to combine vectors by rows (rbind) or columns (cbind)
```{r Matrices 2}
m1<-rbind(vec1,vec2)
m2<-cbind(vec3,vec4)
```

Subsetting from a ``` matrix```. Note that here we use two values for subsetting.
```{r Matrices 3}

# All elements in first row of m1
m1[1,]

#All elements in first column of m1

m1[,1]

```
<span style="color:orangered">**Practice**:</span> Subset from m2
```{r Matrices 4}


# All elements in second row


#All elements in third column





```

###<span style="color:cadetblue">Lists</span>
```Lists``` are really containers for vectors and matrices. They can hold different data classes in each entry of the list including lists! This means you can make lists of lists, which is why lists are considered a recursive vector.

*Lists of lists of lists can confusing*

```{r Lists}
#make a list
w <- list(name="Odum", mynumbers=vec1, mymatrix=m, age=5.3)
primes <- list(
    2, 3, 4:5,
    6:7, 8:11, 12:13,
    14:17, 18:19, 20:23)
#check the class of an entry

```

###<span style="color:cadetblue">Factors</span>
```Factors``` are used for ordered or unordered categorical data. This data type is important for modeling and plotting work. Factors can only contain pre-defined values and are really a way at add labels to integers. 

order categories: `low`,`med`, `high`
unorder categories: `blue`, `red`

```{r Factors1}
dat<-c('low','med', 'high')

# check if it is a factor
print(is.factor(dat))

#Making the vector a into a factor

factor_dat <- factor(dat)

#check again


#Changing the order
new_order_dat <- factor(factor_dat,levels = c("med","high","low"))
```

<span style="color:orangered">**Practice**:</span> Making a ```vector``` into a ```factor```
```{r Factors2}

hues <- c(rep("blue",20), rep("red", 30)) 
# make hues a factor

#summarize hues
summary(hues)
```


###<span style="color:cadetblue">Data frames</span>
```Data frames``` are the most common way we organize ecological data. They have the freedom of lists and the organization of a matrix. Each column can be a different data class. This means for a single observation or row in a data frame categorical, numeric, factorial, etc. can be included.  

```{r Data frames}
# creating vectors
 n = c(2, 3, 5) 
 s = c("aa", "bb", "cc") 
 b = c(TRUE, FALSE, TRUE) 
 # creating the data frame
 df = data.frame(n, s, b)
 
names(df) <- c("ID","Code","Passed") # variable names
```

Like with the ```matrix``` we can subset a ```data frame```
```{r Data frames1}
# subsetting a column
df[,1]

# subsetting a row
df[1,] 

# the $ for subsetting

df$n
```

##<span style="color:cadetblue"> Useful functions</span>

    head() - see first 6 rows
    tail() - see last 6 rows
    dim() - see dimensions
    nrow() - number of rows
    ncol() - number of columns
    str() - structure of each column
    names() - will list the names attribute for a data frame (or any object really), which gives the column names.

## <span style="color:cadetblue">Going further</span>

Tidyverse tibble, etc...

http://www.statmethods.net/input/datatypes.html

