---
title: "M3 Manipulating Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

# 3. Manipulating Data

This module will cover...

data structure
assigning objects
plot function in base R
read in data
logic statements
calculate simple stats in base R


This module should be completed as a team of two, a driver and navigator. The driver will type the code, while the navigator checks for errors, and suggests answers. A good navigator will troubleshoot syntax questions by reading manual pages, googling, etc. 


## Loading data

Before we jump into calculating summary statistics and simple data vizualization, we need to import or load data into `R`. The data that we'll use for this exercise was emailed to you this morning. Save the `.csv` file to your working directory.

The `.csv` contains responses from the Cohort survey we sent out earlier in the week. Let's take a quick look at the file outside of R. What are the different columns and rows? What type of data structure would this be in `R`? 

**Insert screenshot of csv**

A common way to read data frames into R uses `read.csv`. *set up realistic part of example* Work with your partner to complete the `manipulating_data.R` script. This script follows a realiztic work flow used to explore new data. Data exploration might include answering the following:
   
   1.  Exploring Data Structure
      * How many people responded to the survey?
      * What are the titles for the different columns?
      * etc.
   2.  Exploring Data with Summary Statistics  
      * What is the max, min and mean distance to the students favplace? 
      * What is the max, min and mean for all the columns?  
      * Other descriptive statistics built into `R` include: mean, sd, var, min, max, median, range, and quantile
    
  3. Exploring Data with Basic Vizualizations
      * scatter plot: Does birth order correlate with number of siblings? 
      * histogram: What is the distribution of the no. of siblings in the cohort?
      * boxplot: How does the distribution of no. of siblings, birth order and birth date relate?
   

## Solutions to `manipulating_data.R` 

```{r read in csv}


```



```{r exploring data structure}

```




```{r summary statistics}


```



```{r basic vizualizations}


```

