---
title: ""
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

#<span style="color:cadetblue">Manipulating Data</span>

***

This module will cover...

data structure
assigning objects
plot function in base R
read in data
logic statements
calculate simple stats in base R


This module should be completed as a team of two, a driver and navigator. The driver will type the code, while the navigator checks for errors, and suggests answers. A good navigator will troubleshoot syntax questions by reading manual pages, googling, etc. 


## Loading data

Before we jump into calculating summary statistics and simple data vizualization, we need to import or load data into `R`. The data that we'll use for this exercise was emailed to you this morning. Save the `.csv` file to your working directory.

The `.csv` contains responses from the Cohort survey we sent out earlier in the week. Let's take a quick look at the file outside of R. What are the different columns and rows? What type of data structure would this be in `R`? 

**Insert screenshot of csv**

A common way to read data frames into R uses `read.csv`. *set up realistic part of example* Work with your partner to complete the `manipulating_data.R` script. This script follows a realiztic work flow used to explore new data. Data exploration might include answering the following:
   
   1.  Exploring Data Structure
      * How many people responded to the survey?
      * What are the titles for the different columns?
      * etc.
   2.  Exploring Data with Summary Statistics  
      * What is the max, min and mean distance to the students favplace? 
      * What is the max, min and mean for all the columns?  
      * Other descriptive statistics built into `R` include: mean, sd, var, min, max, median, range, and quantile
    
  3. Exploring Data with Basic Vizualizations
      * scatter plot: Does birth order correlate with number of siblings? 
      * histogram: What is the distribution of the no. of siblings in the cohort?
      * boxplot: How does the distribution of no. of siblings, birth order and birth date relate?
   
***

## Solutions to `manipulating_data.R` 

```{r read in csv}
#load survey results into an object called cohort
cohort <- read.csv("CohortData.csv") #the file needs to be located in the working directory  
```


###1. Exploring Data Structure


```{r exploring data structure}
#What is the structure of the data?
str(cohort) #object names are case sensitive 

#What are the names of the different columns?
colnames(cohort) #searching ??colname or tabing out will lead to the command name "colnames"

#How many people responsed to the survey?
#a. modifying the arguement given to length
length(cohort$sibling) #length only works on a ve

#b. using the function dim()
dim(cohort) #dimensions (rows, col) 

#c. using the function nrow()
nrow(cohort) #returns number of rows of a vector, array or data frame
```

###2. Exploring Data with Summary Statistics


```{r summary statistics}

#What is the max, min and mean distance to the students favplace? 
mean(cohort$favplace) #return the mean distance
min(cohort$place) #return the min distance
max(cohort$favplace) #return the max distance

#Use the apply function to calculate the mean, min and max for all the interger columns 
apply(cohort[,1:4], 2, max) #the third arguement is the function
apply(cohort[,1:4], 2, min)
apply(cohort[,1:4], 2, range) #range will return the min and max
apply(cohort[,1:4], 2, mean)


#Use the summary function to create summary statistics for all of the columns
summary(cohort) #returns min,max, median, and quartiles by column


```


###3. Exploring Data with Basic Visualizations


```{r basic vizualizations}

#What is the distribution of the no. of siblings in the cohort?
hist(cohort$sibling) #must call the col with the data of interest
  
#Does birth order correlate with no. of siblings?
plot(y=cohort$sibling, x=cohort$birth.order,ylim=c(0,10), col="black", pch=19, xlab="birth order", ylab="no. of siblings") #plot needs x and y values

#How do the quartiles of the different variables relate to eachother?
boxplot(cohort)

#How does the distribution of no. of siblings, birth order and birth date relate?
boxplot(cohort[,1:3]) #the subset using [rows, col] 


```

