---
title:  <span style="color:cadetblue">Writing a script</span>
author: "A Bento & R Kaul"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#<span style="color:cadetblue">The importance of a script</span>
>1. Permanent 
>2. Repeatable
>3. Annotated 
>4. Shareable

We will do this exercise together...

##<span style="color:cadetblue">Step 1</span>

```
File --> New File --> R Script or Top left green and white plus sign drop down then select R Script
```

##<span style="color:cadetblue">Step 2</span>

``` 
The # symbols declares that anything after it is a comment

You should start with some information at the top, for example:
#############################
## Author:                ###
## Date:                  ###
## Purpose:               ###
#############################
```

##<span style="color:cadetblue">Step 3</span>

Leave an empty "blank line" - These are important  to create chunks of code and annotation

##<span style="color:cadetblue">Step 4</span>

``` 
Next we recomend you always start a script by clearing R's brain
rm(list=ls())
```

##<span style="color:cadetblue">Step 5</span>

Leave another empty "blank line"

##<span style="color:cadetblue">Step 6</span>

Now you are ready to start writing your first script!

```{r loop}
 for(i in 1:10){ # for loop run from 1 to 10
print("Hello world!") # write Hello world 10 times
print(i*i) # also print each number multiplied by itself 
}
```

<span style="color:orangered">**_Check_**:</span> Notice the colours, this is syntax highlighting. Separates:
 
>1. comments
>2. R functions
>3. numbers 
>4. other things


You will notice Rstudio **closes brackets** for you automatically

##<span style="color:cadetblue">Step 7</span>

Once you are done writing your script and you are happy with the commenting. **Save** it in your **working directory** of choice. Give it a meaningful name.


##<span style="color:cadetblue">Step 8</span>

Open your script and run it!
```
setwd("your path") # sets your working directory. You can also do this mannually
source("Script1.R") # source the script of your choice
```
<span style="color:orangered">**_Or_**</span> you can double click on your R script open it in R studio and run specific parts of it
``` 
Mac: cmd+return 
Windows: ctrl+enter
or pressing Run 
```

