---
title: "Data Visualization: Group 2"
author: "Ana Bento & Reni Kaul"
date: "August 18-19 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```



Yesterday we started to explore data visualization in base `R`. Today we will spend time exploring different visualizations using the `ggplot2` package. The different approaches each have their advantages, so you will likely use both during your research career. 



#Problem 1

  Yesterday we explored the `nycflights13` dataset. One plot was a boxplot of the temperature by month using base `R`. The plot looked like this:
  
```{r boxplot}
library(nycflights13) #load data from package

boxplot(weather$temp ~ weather$month, xlab="month", ylab="Temperature (F)" ) #instead of assigning x and y values, you can use the syntax y ~ x
```


Today, your goal is to recreate this figure using the `ggplot2` package.  `ggplot2` was first published to CRAN in 2005, so a quick google search will return *MANY* tutorials, discussion pages, etc.  The general syntax for boxplots in `ggplot2` is:

```{r function structure, eval=FALSE}
ggplot(data, #matrix
       aes(x= X.col, y= Y.col)) + #x and y columns in matrix to plot
      geom_boxplot() #plot the data as a box plot
```


A basic `ggplot2` boxplot should look like this:

```{r gg boxplot, echo=FALSE }
library(ggplot2)
weather$Month <- factor(weather$month,
                           labels = c("Jan","Feb","March", "April", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
ggplot(weather, aes(x=as.factor(month), y=temp)) + geom_boxplot()

```

With additional customization, we can produce a plot with a little more shine. For example, this plot has transparent colored boxes and month names instead of numbers. 

```{r gg boxplot2, echo=FALSE}
fill <- "#4271AE"
line <- "#1F3552"
weather$Month <- factor(weather$month,
                           labels = c("Jan","Feb","March", "April", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
ggplot(weather, aes(x=Month, y=temp)) + 
  geom_boxplot(fill=fill, line=line, alpha=0.5) +
  scale_y_continuous(name="Temperature") +
  ggtitle("Temperature in NYC 2013")

```

Work with your team to create a polished `ggplot2` boxplot. It doesn't need to look like these figures, but you are welcome to use them as a guide. A very detailed tutorial is on [ZevRoss](http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/). A more formal guide is in the [R for Data Science](http://r4ds.had.co.nz/data-visualisation.html) visualization chapter. 

#Problem 2

For this exercise, we will use `diamonds` in the `ggplot2` package. The `ggplot2` package should already be loaded. Explore the `diamonds` dataset which contains 10 attributes (price, cut, clarity, etc.) for ~54,000 diamonds. 

So far we have only created single panel visualizations. Making multipanel visualizations is a quick way to look at trends between covariates. Here the x-axis is diamond carat, y-axis price, and each panel is diamond cut.  

```{r ggscatter, echo=FALSE }
ggplot(diamonds) +
  geom_point(aes(carat, price), alpha = 0.1) +
  facet_wrap(~cut, ncol = 3)
```


Your goal is to create this figure, and modify it in at least 2 ways. The syntax for a multipanel or facet plot is similar to a scatterplot but uses `facet_wrap()` in addition to `geom_point()`. 
