---
title: 'Data Visualization: Group 3'
author: "Ana Bento & Reni Kaul"
date: "August 18-19 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```




Yesterday we started to explore data visualization in base `R`. Today we will spend time exploring different visualizations using the `ggplot2` package. The different approaches each have their advantages, so you will likely use both during your research career. 



#Problem 1

  Yesterday we explored the `nycflights13` dataset. One plot was a boxplot of the temperature by month using base `R`. The plot looked like this:
  
```{r boxplot}
library(nycflights13) #load data from package

boxplot(weather$temp ~ weather$month, xlab="month", ylab="Temperature (F)" ) #instead of assigning x and y values, you can use the syntax y ~ x
```


Today, your goal is to recreate this figure using the `ggplot2` package.  `ggplot2` was first published to CRAN in 2005, so a quick google search will return *MANY* tutorials, discussion pages, etc.  The general syntax for boxplots in `ggplot2` is:

```{r function structure, eval=FALSE}
ggplot(data, #matrix
       aes(x= X.col, y= Y.col)) + #x and y columns in matrix to plot
      geom_boxplot() #plot the data as a box plot
```


A basic `ggplot2` boxplot should look like this:

```{r gg boxplot, echo=FALSE }
library(ggplot2)
weather$Month <- factor(weather$month,
                           labels = c("Jan","Feb","March", "April", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
ggplot(weather, aes(x=as.factor(month), y=temp)) + geom_boxplot()

```

With additional customization, we can produce a plot with a little more shine. For example, this plot has transparent colored boxes and month names instead of numbers. 

```{r gg boxplot2, echo=FALSE}
fill <- "#4271AE"
line <- "#1F3552"
weather$Month <- factor(weather$month,
                           labels = c("Jan","Feb","March", "April", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
ggplot(weather, aes(x=Month, y=temp)) + 
  geom_boxplot(fill=fill, line=line, alpha=0.5) +
  scale_y_continuous(name="Temperature") +
  ggtitle("Temperature in NYC 2013")

```

Work with your team to create a polished `ggplot2` boxplot. It doesn't need to look like these figures, but you are welcome to use them as a guide. A very detailed tutorial is on [ZevRoss](http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/). A more formal guide is in the [R for Data Science](http://r4ds.had.co.nz/data-visualisation.html) visualization chapter. 

#Problem 2

The `ggplot2` package can make highly stylized visualizations with relatively few codes of line. The violin plot is a good example of these stylized visualization. [Violin plots](http://www.letmegooglethat.com/?q=violin+plots) are a combination of a histogram and boxplot. Today you'll create a violin plot with data pulled from a unimodal, bimodal, and trimodal distribution. 

```{r}
###Sample gaussian distributions with 1, 2 and 3 modes.
df <- data.frame(
  "Distribution" = c(rep("Unimodal", 500),
                     rep("Bimodal", 250),
                     rep("Trimodal", 600)),
  "Value" = c(rnorm(500, 6, 1),
              rnorm(200, 3, .7), rnorm(50, 7, 0.4),
              rnorm(200, 2, 0.7), rnorm(300, 5.5, 0.4), rnorm(100, 8, 0.4))
)

# Reorder levels, this will be needed for the plotting
df$Distribution <- factor(df$Distribution,
                          levels(df$Distribution)[c(3, 1, 2)])

```

This plot was created from the `df` object produced above. 

```{r, echo=FALSE}
ggplot(df, aes(Distribution, Value)) + geom_violin(aes(fill = Distribution))

```

Your goal is to create this figure, and modify it in 2 ways. The syntax for a scatterplot is similar to a boxplot but `geom_violin()` is used in place of `geom_boxplot()`. 
